### 2012-01-29 docs/getting_started/getting_started.tex
more detailed quick start, including how to get boxlib

@@

Different problems in \maestro\ are contained in individual
sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
in each problem directory lists the components of {\tt MAESTRO}
that are used to build the executable.  Full details of the
{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.

@@

\section{Compiling and Running}

To build the \maestro\ executable, simply type `{\tt make}' in the
problem directory.  The executable will have a name like {\tt
  main.Linux.Intel.exe}, where the specific parts of the name depend
on the compilers and OS used.  \maestro\ can be built with support for
parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
The top portion of the {\tt GNUmakefile} can be edited to choose the
build options.  Setting {\tt MPI := t} will build the code to use MPI
for parallelization.  Setting {\tt OMP := t} will use OpenMP.  OpenMP
is used together with MPI, with MPI distributing the grids across the
processors and within a shared-memory node, OpenMP allows many cores
to operate on the same grid.  A final build option determines whether
to build the code with support for debugging.  Setting {\tt NDEBUG :=
  t} disables debugging features.

If the {\tt helmeos} equation of state is chosen for the problem, then
the data table for the equation of state, {\tt
  extern/EOS/helmeos/helm\_table.dat} needs to be copied to the
directory in which the code is run.

Each problem will have one or more inputs files (for example, {\tt
  test2/inputs\_2d}) which override the default values for one or more
runtime parameters.  The code is run simply as:
\begin{verbatim}
  ./main.Linux.Intel.exe inputs_2d
\end{verbatim}
for a serial executable, and, depending on your MPI environment, something like:
\begin{verbatim}
  mpiexec -n 16 ./main.Linux.Intel.mpi.exe inputs_2d
\end{verbatim}
for the parallel code.

As the code runs, it will output both plotfiles and checkpoints as
well as one of more text diagnostic files (like {\tt
  maestro\_diag.out}) with integral or extrema information (like
maximum Mach number) from each timestep.  By default, the plotfiles
will be named {\tt plt}{\em nnnnn}, where the number {\em nnnnn} is
the timestep number when the file was outputted.  Similarly, the
checkpoints are named {\tt chk}{\em nnnnn}.  BoxLib plotfiles and
checkpoints are actually directories, with the data stored in
sub-directories grouped by refinement level.  Details of the simulation
(build information, number of processors used, output date, output
directory, runtime parameter values, ...) are stored in the {\tt
  job\_info} file in each plotfile directory.

A large number of runtime parameters affect the code's behavior.  Any
runtime parameter can be set either in the inputs file, as {\tt parameter
  = value}, or on the command line by adding {\tt --parameter value} after
listing the inputs file on the command line.  A full list of runtime
parameters is detailed in \S~\ref{sec:runtime_parameters}.  Below we
outline the most common runtime parameters.

\section{Common Runtime Parameters}

@@

Visualization and analysis are done with on the plotfiles.  A
number of in-house and externally developed tools can work 
with BoxLib-formatted plotfiles.

VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\
directly from BoxLib plotfiles (to open a plotfile in VisIt, point to
the {\tt Header} file inside the plotfile directory).  Details on
visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.

>>>>

%-----------------------------------------------------------------------------
% Quick Start
%-----------------------------------------------------------------------------
\section{Quick Start}

Here we will run the standard {\tt test2} problem (three reacting
bubbles in a plane-parallel stratified atmosphere) on a single
processor.  This test problem was shown in paper~3.

\begin{enumerate}

\item Get a copy of \boxlib.

\maestro\ requires the \boxlib\ library to manage the grids and 
parallelization.  We also rely on the build system in \boxlib\ to
build a \maestro\ executable.  \boxlib can be obtained via {\tt
git}\footnote{{\tt git} is an open-source source code version control
system.  It is a standard part of any Linux distribution.  You may
need to use your system's package manager to install {\tt git} if it
is not already present} as:
\begin{verbatim}
git clone https://ccse.lbl.gov/pub/Downloads/BoxLib.git
\end{verbatim}

\item Setup your shell environment.

\maestro\ needs to know where to find \boxlib\, by specifying the 
{\tt BOXLIB\_HOME} environment variable.

If you shell is {\tt Bash}, add
\begin{verbatim}
export BOXLIB_HOME="/path/to/BoxLib/"
\end{verbatim}
to your {\tt .bashrc}.

If your shell is{\tt Csh/Tcsh}, add
\begin{verbatim}
setenv BOXLIB_HOME /path/to/BoxLib/
\end{verbatim}
to your {\tt .cshrc}.

\item Setup the problem's {\tt GNUmakefile}.

In MAESTRO, each problem gets their own sub-directory under {\tt
MAESTRO/}.  This sub-directory will contain any problem-specific files
as well as the {\tt GNUmakefile} that specifies how to build the executable.
Note: we rely on features of GNU {\tt make}.  Full details of the
{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.  Here
we will configure for a serial build.

We only need to worry about the options at the very top of the
{\tt GNUmakefile} for now.  These should be set as follows:
\begin{itemize}
\item {\tt NDEBUG := t}

This option determines whether we compile with support for debugging
(usually also enabling some runtime checks).  Setting {\tt NDEBUG := t}
turns off debugging.

\item {\tt MPI := }

This determines whether we are doing a parallel build, using the Message
Passing Interface (MPI) library.  For now, we leave this option empty,
disabling MPI.  This will build \maestro\ in serial mode, so no MPI
library needs to be present on the host system.

\item {\tt OMP := }

This determines whether we are using OpenMP to do parallelism within a
shared memory node.  OpenMP is used together with MPI, with MPI
distributing the grids across the processors and within a
shared-memory node, OpenMP allows many cores to operate on the same
grid.  For now, we leave this option empty, disabling OpenMP.

\item {\tt SDC := }

This option determines whether we want to use the alternate SDC
(Spectral Deferred Corrections) source for \maestro\ (see
Chapter \ref{ch:sdc}).  This is experimental, and this option not present
in all problems.  We leave this blank so we compile the
default \maestro\ source.

\item {\tt COMP := gfortran}

This option specifies the Fortran compiler.  We will use {\tt gfortran}, which is
the preferred compiler for \maestro.  Specifying this compiler will automatically
pull in the compiler settings as specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.

\item Build the executable.

Type {\tt make}.  The build system will first find the dependencies
amongst all the source files and then build the executable.  When
finished, the executable will have a name like {\tt
main.Linux.gfortran.exe}, where the specific parts of the name depend
on the compilers and OS used.

\item Copy in auxillary files.

This problem uses the {\tt helmeos} equation of state, so 
we need to copy in the data table for the equation of state, {\tt
  extern/EOS/helmeos/helm\_table.dat}.

\begin{verbatim}
cp ../../extern/EOS/helmeos/helm_table.dat .
\end{verbatim}

\item Run.

Each problem requires an input file.  The inputs file consists of
lines of the form {\em parameter = value}, where {\em parameter} is
one of the many runtime parameters \maestro\ knows, and {\em value}
overrides the default value for that parameter.  For the {\tt test2}
problem, we will use the inputs file {\tt inputs\_2d}.  An overview of
some of the more common runtime parameters is given
in \S~\ref{sec:gettingstarted:runtime}, and a full list of
all \maestro\ runtime parameters and their default values is given in
Chapter~\ref{ch:runtimeparameters}.

\maestro\ is run simply as:
runtime parameters.  The code is run simply as:
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d
\end{verbatim}
We can also override the default value of any runtime parameter by specifying
it on the commandline as
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d --parameter value
\end{verbatim}

As the code runs, a lot of information will pass through the screen.
For each timestep, each of the steps 1 through 12 shown in
the \maestro\ flowchart (Chapter~\ref{ch:flowchart}) will be shown
along with diagnostic information about the solution.  Upon completion
some memory usage information is printed.

\item Examine the output.

As the code runs, it will output both plotfiles and checkpoints as
well as one of more text diagnostic files ({\tt maestro\_diag.out} by
default) with integral or extrema information (like maximum Mach
number) from each timestep.

By default, the plotfiles will be named {\tt plt}{\em nnnnn}, where
the number {\em nnnnn} is the timestep number when the file was
outputted.  Similarly, the checkpoints are named {\tt chk}{\em
nnnnn}.  \boxlib\ plotfiles and checkpoints are actually directories,
with the data stored in sub-directories grouped by refinement level.
Details of the simulation (build information, number of processors
used, output date, output directory, runtime parameter values, ...)
are stored in the {\tt job\_info} file in each plotfile directory.

Visualization of results is described in the next section.

\end{enumerate}

Visualization and analysis are done with on the plotfiles.  A
number of in-house and externally developed tools can work 
with \boxlib-formatted plotfiles.

VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\
directly from \boxlib\ plotfiles (to open a plotfile in VisIt, point to
the {\tt Header} file inside the plotfile directory).  Details on
visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.

@@

Different problems in \maestro\ are contained in individual
sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
in each problem directory lists the components of {\tt MAESTRO}
that are used to build the executable.

@@

\section{Common Runtime Parameters}
\label{sec:gettingstarted:runtime}

@@
